# Hello and welcome

This document is meant as a review / for improving my own understanding of statistics. Use it if you can but I take no responsibility for what is right and wrong. It is meant to be very hands on with code. Most often only the output is shown, but if you want to see it it is all available on my github page. 

## How to Read These Notes

This collection of statistics notes is organized to help you understanding statistical concepts, from foundational theory to practical applications.

### Structure

The notes are organized into four main sections, accessible through the navigation bar at the top:

1. **Introduction** - What is statistics? Probability and statistical inference
2. **Distributions** - Discrete and continuous probability distributions
3. **Statistical Inference** - Frequentist and Bayesian statistics, hypothesis testing
4. **Machine Learning** - Statistical learning methods and deep learning

### Content Types

These notes use a mixture of descriptive text, mathematical formulas, figures, and interactive elements (R code, visualizations, and Shiny apps).

#### Code Examples

Some code is placed in the margin for reference:

:::{.column-margin}
```{r echo=True}
#| echo: false
file.path(R.home("bin"), "R")
```
:::

:::{.column-margin}
```{python}
#| echo: false
import os
import sys
print(sys.executable)
```
:::

Some code is embedded in the main panel:

```{r}
1+1
```

```{python}
1+1
```

#### Figures

Figures are included in the main text:

![](illustrations/figures/example.svg)

Or in the margin:

:::{.column-margin}
![](illustrations/figures/example.svg)
:::

### Suggested Reading Order

For a comprehensive understanding, I recommend starting with the **Introduction** section and then moving through the sections from left to right. However, feel free to explore topics in any order based on your interests and needs.