# Beta Distribution

## Overview

The beta distribution is a continuous probability distribution defined on the interval [0, 1], which makes it ideal for modeling probabilities and proportions.

## Description

The beta distribution is parameterized by two positive shape parameters, $\alpha$ and $\beta$, which control the shape of the distribution. It can take on a variety of shapes depending on these parameters:

- $\alpha < 1, \beta < 1$: U-shaped
- $\alpha = \beta$: Symmetric
- $\alpha > 1, \beta > 1$: Unimodal (single peak)
- $\alpha = 1, \beta = 1$: Uniform distribution on [0, 1]

The beta distribution is commonly used as a **prior distribution** in Bayesian statistics, particularly for modeling probabilities and rates.

## Probability Density Function

$$f(x | \alpha, \beta) = \frac{x^{\alpha - 1} (1 - x)^{\beta - 1}}{B(\alpha, \beta)} \quad \text{for } 0 \leq x \leq 1$$

Where:
- $\alpha$ = first shape parameter (often called "alpha")
- $\beta$ = second shape parameter (often called "beta")
- $B(\alpha, \beta)$ = Beta function, which normalizes the distribution
- The Beta function is: $B(\alpha, \beta) = \frac{\Gamma(\alpha)\Gamma(\beta)}{\Gamma(\alpha + \beta)}$

## Properties

- Mean = $\frac{\alpha}{\alpha + \beta}$
- Mode = $\frac{\alpha - 1}{\alpha + \beta - 2}$ (if $\alpha > 1$ and $\beta > 1$)
- Variance = $\frac{\alpha \beta}{(\alpha + \beta)^2 (\alpha + \beta + 1)}$

## Visualization

```{r}
#| echo: false
library('tidyverse')
alpha_vals <- c(0.5, 2, 5)
beta_vals <- c(0.5, 2, 5)
x <- seq(0, 1, length.out = 500)

# Create a grid of plots
tibble(alpha = rep(alpha_vals, each = 500), 
       beta = rep(beta_vals, times = 1500),
       x = x) %>
  mutate(y = dbeta(x, shape1 = alpha, shape2 = beta)) %>
  ggplot(aes(x = x, y = y, color = factor(alpha) + " / " + factor(beta))) +
  geom_line(size = 1) +
  facet_grid(beta ~ alpha) +
  labs(x = 'x', y = 'Density', title = 'Beta Distribution for various α, β') +
  theme_minimal(base_size = 14) +
  theme(legend.position = 'none')
```

## Interactive App

```{r}
#| echo: false
#| error: false
if (requireNamespace('shiny', quietly = TRUE) && file.exists('/home/jenswaaben/phd/courses/stats_personal/statistics/apps/continuous_distributions/beta_app/app.R')) {
  shiny::shinyAppFile('/home/jenswaaben/phd/courses/stats_personal/statistics/apps/continuous_distributions/beta_app/app.R')
}
```

## Real-world Examples

- Prior distribution for a probability in Bayesian analysis
- Modeling the proportion of time spent on different activities
- Distribution of allele frequencies in population genetics
- Proportions of voters supporting different candidates
- Probabilities of success in various contexts
