# Chi-Squared Distribution

## Overview

The chi-squared distribution (also written as χ² distribution) is one of the most fundamental distributions in statistical inference.

## Description

If $Z_1, Z_2, ..., Z_k$ are independent standard normal random variables (mean = 0, variance = 1), then the sum of their squares follows a chi-squared distribution with k degrees of freedom:

$$Q = Z_1^2 + Z_2^2 + \cdots + Z_k^2 \sim \chi^2_k$$

The chi-squared distribution is widely used in hypothesis testing, particularly in chi-squared tests for goodness-of-fit and independence.

## Probability Density Function

$$f(x | k) = \frac{x^{k/2 - 1} e^{-x/2}}{2^{k/2} \Gamma(k/2)} \quad \text{for } x \geq 0$$

Where:
- $k$ = degrees of freedom (positive integer)
- $\Gamma(\cdot)$ = Gamma function

## Properties

- Defined on [0, ∞)
- Mean = $k$
- Variance = $2k$
- Mode = $k - 2$ (for k ≥ 2)
- Median ≈ $k - 2/3$ for large k

## Visualization

```{r}
#| echo: false
library('tidyverse')
df_vals <- c(1, 2, 5, 10)
x <- seq(0, 30, length.out = 500)

# Calculate densities for each df
results <- map_dfr(df_vals, ~ {
  tibble(x = x, y = dchisq(x, df = .x), df = .x)
})

results %>
  ggplot(aes(x = x, y = y, color = factor(df))) +
  geom_line(size = 1) +
  labs(x = 'x', y = 'Density', title = 'Chi-Squared Distribution for various degrees of freedom') +
  theme_minimal(base_size = 14)
```

## Interactive App

```{r}
#| echo: false
#| error: false
if (requireNamespace('shiny', quietly = TRUE) && file.exists('/home/jenswaaben/phd/courses/stats_personal/statistics/apps/continuous_distributions/chi_squared_app/app.R')) {
  shiny::shinyAppFile('/home/jenswaaben/phd/courses/stats_personal/statistics/apps/continuous_distributions/chi_squared_app/app.R')
}
```

## Real-world Examples

- Chi-squared goodness-of-fit test
- Chi-squared test of independence
- Confidence intervals for variance
- Likelihood ratio tests
- Analysis of variance (ANOVA)
