# Gamma Distribution

## Overview

The gamma distribution is a **continuous probability distribution** that generalizes the exponential distribution and arises naturally in processes with waiting times for multiple events.

## Description

The gamma distribution is often used to model the time until a specified number of events occur in a Poisson process. While the exponential distribution models the time until the first event, the gamma distribution models the time until the k-th event.

Formally, if $X_1, X_2, ..., X_k$ are independent and identically distributed exponential random variables with rate parameter $\lambda$, then their sum follows a gamma distribution:

$$Y = X_1 + X_2 + \cdots + X_k \sim \text{Gamma}(k, \lambda)$$

## Probability Density Function

There are two common parameterizations of the gamma distribution:

### Shape-Rate Parameterization

$$f(x | k, \theta) = \frac{\theta^k x^{k - 1} e^{-\theta x}}{\Gamma(k)} \quad \text{for } x > 0$$

### Shape-Scale Parameterization (used by R)

$$f(x | \alpha, \beta) = \frac{x^{\alpha - 1} e^{-x / \beta}}{\beta^\alpha \Gamma(\alpha)} \quad \text{for } x > 0$$

Where:
- $\alpha$ (alpha) or $k$ = shape parameter (> 0)
- $\beta$ (beta) or $\theta$ = scale parameter (or its inverse, the rate parameter) (> 0)
- $\Gamma(\alpha)$ = Gamma function, which generalizes the factorial

Note: In R, `dgamma(x, shape, scale)` uses the shape-scale parameterization where `rate = 1/scale`.

## Properties

- Defined on (0, ∞)
- Mean = $\alpha \beta$ (shape-scale) or $k / \theta$ (shape-rate)
- Median ≈ $\alpha \beta (1 - 1/(9\alpha))^3$ for large $\alpha$
- Mode = $(\alpha - 1) \beta$ for $\alpha \geq 1$ (shape-scale)
- Variance = $\alpha \beta^2$ (shape-scale) or $k / \theta^2$ (shape-rate)

## Special Cases

- Exponential distribution: $\alpha = 1$ (shape parameter = 1)
- Chi-squared distribution: $\beta = 2$, $\alpha = k/2$ where k is degrees of freedom
- Erlang distribution: $\alpha$ is a positive integer

## Visualization

```{r}
#| echo: false
library('tidyverse')
shape_vals <- c(1, 2, 5)
scale <- 2
x <- seq(0, 20, length.out = 500)

# Calculate densities for each shape
results <- map_dfr(shape_vals, ~ {
  tibble(x = x, y = dgamma(x, shape = .x, scale = scale), shape = .x)
})

results %>
  ggplot(aes(x = x, y = y, color = factor(shape))) +
  geom_line(size = 1) +
  labs(x = 'x', y = 'Density', 
       title = paste('Gamma Distribution (scale =', scale, ')', sep = '')) +
  theme_minimal(base_size = 14)
```

## Interactive App

```{r}
#| echo: false
#| error: false
if (requireNamespace('shiny', quietly = TRUE) && file.exists('/home/jenswaaben/phd/courses/stats_personal/statistics/apps/continuous_distributions/gamma_app/app.R')) {
  shiny::shinyAppFile('/home/jenswaaben/phd/courses/stats_personal/statistics/apps/continuous_distributions/gamma_app/app.R')
}
```

## Real-world Examples

- Time until the k-th customer arrives at a store
- Time until the k-th machine component fails
- Time until the k-th earthquake in a region
- Rainfall accumulation over time
- Service times in queuing systems
- Survival analysis
- Financial modeling (time until default)
- Reliability engineering
