# Normal Distribution

## Overview

The normal distribution, also known as the **Gaussian distribution** or **bell curve**, is perhaps the most important and well-known probability distribution in statistics.

## Description

It is symmetric around its mean, with values closer to the mean being more likely than values farther away. The shape of the distribution is determined by its mean ($\mu$) and standard deviation ($\sigma$).

Many natural phenomena follow (approximately) a normal distribution due to the Central Limit Theorem, which states that the sum of many independent random variables tends to be normally distributed, regardless of their underlying distributions.

## Probability Density Function

$$f(x | \mu, \sigma^2) = \frac{1}{\sqrt{2\pi\sigma^2}} \exp\left(-\frac{(x - \mu)^2}{2\sigma^2}\right)$$

Where:
- $\mu$ = mean
- $\sigma^2$ = variance
- $\sigma$ = standard deviation

## Properties

- Mean = $\mu$
- Median = $\mu$
- Mode = $\mu$
- Variance = $\sigma^2$
- Symmetric about the mean
- Approximately 68% of data within $\mu \pm \sigma$
- Approximately 95% of data within $\mu \pm 2\sigma$
- Approximately 99.7% of data within $\mu \pm 3\sigma$

## Visualization

```{r}
#| echo: false
library('tidyverse')
mean_val <- 0
sd_val <- 1
x <- seq(mean_val - 4*sd_val, mean_val + 4*sd_val, length.out = 500)
y <- dnorm(x, mean = mean_val, sd = sd_val)
tibble(x = x, y = y) %>
  ggplot(aes(x = x, y = y)) +
  geom_line(color = '#3498db', size = 1.5) +
  geom_vline(xintercept = mean_val, color = 'red', linetype = 'dashed') +
  geom_hline(yintercept = 0, color = 'black') +
  labs(x = 'x', y = 'Density',
       title = paste('Normal Distribution (μ=', mean_val, ', σ=', sd_val, ')', sep = '')) +
  theme_minimal(base_size = 16)
```

## Interactive App

```{r}
#| echo: false
#| error: false
if (requireNamespace('shiny', quietly = TRUE) && file.exists('/home/jenswaaben/phd/courses/stats_personal/statistics/apps/continuous_distributions/normal_app/app.R')) {
  shiny::shinyAppFile('/home/jenswaaben/phd/courses/stats_personal/statistics/apps/continuous_distributions/normal_app/app.R')
}
```

## Real-world Examples

- Human height
- Blood pressure
- IQ scores
- Measurement errors
- Many biological measurements
- Financial returns (often approximately normal)
