# Bernoulli Distribution

## Overview

The Bernoulli distribution is the most basic probability distribution, describing a single binary outcome. It is so simple that we often don't even consider it a distribution, yet technically it is.

## Description

The Bernoulli distribution describes two outcomes (e.g., flipping a coin). This is done one time, with a particular probability. It makes the most intuitive sense because distributions are usually considered to describe an array of outputs, not just two values.

For the Bernoulli distribution, the only parameter that changes is **p** (the probability of success).

## Probability Density Function

$$f(k;p) = Prob(X = k) =  p^k(1-p)^{1-k}  \  \  \  for \  k \  in \  \{0,1\}$$

Where:
- $p$ = probability of success (1) / probability of failure (0)
- $k$ = outcome (0 or 1)

## Visualization

```{r}
#| echo: false
library('tidyverse')
tibble(dens = dbinom(c(0,1), 1, 0.5), x = c('Failure', "Success")) %>
      ggplot(mapping = aes(x = x, y = dens, fill = x)) +
        geom_col() +
        ylim(0,1) +
        theme_minimal(base_size = 20) + 
        labs(y = 'Probability')
```

## Interactive App

The Shiny app allows you to explore how changing the probability parameter affects the distribution.

```{r}
#| echo: false
#| error: false
if (requireNamespace('shiny', quietly = TRUE) && file.exists('/home/jenswaaben/phd/courses/stats_personal/statistics/apps/discrete_distributions/bernoulli_app/app.R')) {
  shiny::shinyAppFile('/home/jenswaaben/phd/courses/stats_personal/statistics/apps/discrete_distributions/bernoulli_app/app.R')
}
```

## Real-world Examples

- Coin flip (heads/tails)
- Success/failure of a single trial
- Yes/No questions
- Pass/Fail tests
