# Binomial Distribution

## Overview

The binomial distribution is based on the Bernoulli distribution. It describes the probability of getting a specific number of successes out of N independent Bernoulli trials.

## Description

More specifically, the binomial distribution describes the probability of getting exactly $k$ successes in $n$ independent trials, where each trial has probability $p$ of success.

## Probability Density Function

$$f(k, n, p) = Prob(X = k) = \binom{n}{k}  p^k(1-p)^{n-k}$$

Where:
- $p$ = probability of each trial being a success
- $k$ = number of successes
- $n$ = number of Bernoulli trials
- $\binom{n}{k}$ = binomial coefficient (number of ways to choose k elements from n)

The binomial coefficient can be calculated by:

$$\binom{n}{k} = \frac{n!}{k!(n-k)!}$$

For example: $\binom{4}{2} = 6$ because there are six different ways to choose a subset of two from a set of four different values.

## Visualization

```{r}
#| echo: false
library('tidyverse')
tibble(dens = dbinom(c(0:10), 10, 0.25), x = c(0:10)) %>
      ggplot(mapping = aes(x = x, y = dens)) +
      geom_col(fill = '#528ce1') +
      theme_minimal(base_size = 20) + 
      labs(x = 'Number of successes out of 10 with probability 0.25', 
        y = 'Probability') + 
      scale_x_continuous(breaks = 0:10)
```

## Interactive App

```{r}
#| echo: false
#| error: false
if (requireNamespace('shiny', quietly = TRUE) && file.exists('/home/jenswaaben/phd/courses/stats_personal/statistics/apps/discrete_distributions/binomial_app/app.R')) {
  shiny::shinyAppFile('/home/jenswaaben/phd/courses/stats_personal/statistics/apps/discrete_distributions/binomial_app/app.R')
}
```

## Real-world Examples

- Number of heads in 10 coin flips
- Number of defective items in a batch
- Number of correct answers on a multiple-choice test
- Number of successes in clinical trials
