# Geometric Distribution

## Overview

The geometric distribution describes **how many Bernoulli trials it takes to get a single success**.

## Description

This is the discrete probability distribution of the number of trials needed to get one success, supported on the set {1, 2, 3, ...}. It is a special case of the negative binomial distribution where r = 1.

Note: There are two different parameterizations. This page uses the one that counts the number of trials (including the successful one). Another counts the number of failures before the first success.

## Probability Density Function

$$Pr(X = k) = (1-p)^{k-1} p$$

Where:
- $p$ = probability of success on an individual trial
- $k$ = number of trials until the first success (1, 2, 3, ...)

## Properties

- Mean = $1/p$
- Variance = $(1-p)/p^2$
- The geometric distribution is **memoryless**: the probability of success on the next trial is independent of how many failures have already occurred

## Visualization

```{r}
#| echo: false
library('tidyverse')
p <- 0.2
k_vals <- 1:20
dens <- dgeom(k_vals - 1, prob = p)  # Using R's parameterization (number of failures)
tibble(x = k_vals, y = dens) %>
  ggplot(aes(x = x, y = y)) +
  geom_col(fill = '#8e44ad') +
  labs(x = 'Number of trials until first success', 
       y = 'Probability',
       title = paste('Geometric Distribution (p=', p, ')', sep = '')) +
  theme_minimal(base_size = 16)
```

## Interactive App

```{r}
#| echo: false
#| error: false
if (requireNamespace('shiny', quietly = TRUE) && file.exists('/home/jenswaaben/phd/courses/stats_personal/statistics/apps/discrete_distributions/geometric_app/app.R')) {
  shiny::shinyAppFile('/home/jenswaaben/phd/courses/stats_personal/statistics/apps/discrete_distributions/geometric_app/app.R')
}
```

## Real-world Examples

- Number of coin flips until the first head
- Number of dice rolls until the first six
- Number of job applications until the first interview offer
- Number of dates until finding a compatible partner
- Number of attempts until successfully solving a problem
