# Negative Binomial Distribution

## Overview

Instead of modelling the probability of successes in a certain amount of pulls (like the binomial distribution), the negative binomial distribution models **how many failures we can expect before we get a certain number of successes**.

## Description

This is a discrete probability distribution for the number of failures in a sequence of independent and identically distributed Bernoulli trials before a specified number of successes (denoted $r$) occurs.

## Probability Density Function

$$X \sim NB(r, p)$$

$$f(k; r, p) = Pr(X = k) = \binom{k + r - 1}{k}(1-p)^k p^r$$

Where:
- $k$ = number of failures
- $r$ = number of successes (target count)
- $p$ = probability of success on an individual trial
- The binomial coefficient $\binom{k + r - 1}{k}$ describes in how many different ways k failures can be chosen between k failures and r successes (the minus one is because the last pull is always a success)

## Visualization

```{r}
#| echo: false
library('tidyverse')
# Example visualization of negative binomial
r <- 5
p <- 0.3
k_vals <- 0:20
dens <- dnbinom(k_vals, size = r, prob = p)
tibble(x = k_vals, y = dens) %>
  ggplot(aes(x = x, y = y)) +
  geom_col(fill = '#e74c3c') +
  labs(x = 'Number of failures before 5 successes', 
       y = 'Probability',
       title = paste('Negative Binomial (r=', r, ', p=', p, ')', sep = '')) +
  theme_minimal(base_size = 16)
```

## Interactive App

```{r}
#| echo: false
#| error: false
if (requireNamespace('shiny', quietly = TRUE) && file.exists('/home/jenswaaben/phd/courses/stats_personal/statistics/apps/discrete_distributions/negative_binomial_app/app.R')) {
  shiny::shinyAppFile('/home/jenswaaben/phd/courses/stats_personal/statistics/apps/discrete_distributions/negative_binomial_app/app.R')
}
```

## Real-world Examples

- Number of patients a doctor needs to see before finding 5 with a rare disease
- Number of sales calls needed before closing 10 deals
- Number of games a team needs to play before achieving 3 wins
- Number of manufacturing batches before producing 20 defect-free items
