# Poisson Distribution

## Overview

The Poisson distribution is used extensively in biological and physical scenarios. It was first developed to describe how likely it is to see a certain number of events within a fixed time frame or space.

## Description

For example: how likely is it to see 1, 2, 3 (etc.) cars outside your window in an hour? Or the number of radioactive decays in a given time period? Or the number of calls to a call center per minute?

The Poisson distribution is particularly useful for modeling **rare events** that occur independently at a constant average rate.

## Probability Density Function

$$Pr(X = k) = \frac{\lambda^k e^{-\lambda}}{k!}$$

Where:
- $\lambda$ (lambda) = average number of events in the given time/space interval
- $k$ = number of events (0, 1, 2, ...)
- $e$ = Euler's number (~2.71828)

## Properties

- Mean = $\lambda$
- Variance = $\lambda$
- The distribution is right-skewed for small $\lambda$, becoming more symmetric as $\lambda$ increases

## Visualization

```{r}
#| echo: false
library('tidyverse')
lambda <- 5
k_vals <- 0:15
dens <- dpois(k_vals, lambda = lambda)
tibble(x = k_vals, y = dens) %>
  ggplot(aes(x = x, y = y)) +
  geom_col(fill = '#2ecc71') +
  geom_vline(xintercept = lambda, color = 'red', linetype = 'dashed') +
  labs(x = 'Number of events (k)', 
       y = 'Probability',
       title = paste('Poisson Distribution (λ=', lambda, ')', sep = '')) +
  theme_minimal(base_size = 16)
```

## Interactive App

```{r}
#| echo: false
#| error: false
if (requireNamespace('shiny', quietly = TRUE) && file.exists('/home/jenswaaben/phd/courses/stats_personal/statistics/apps/discrete_distributions/poisson_app/app.R')) {
  shiny::shinyAppFile('/home/jenswaaben/phd/courses/stats_personal/statistics/apps/discrete_distributions/poisson_app/app.R')
}
```

## Real-world Examples

- Number of cars passing a point in an hour
- Number of emails received per day
- Number of radioactive decays per minute
- Number of patients arriving at a hospital emergency room
- Number of typos on a page
- Number of bacteria in a microscope field of view
